\documentclass[12pt]{article}

\input{packages}
\graphicspath{
    {./figures/}
}

%\backgroundsetup{
%  position=current page.east,
%  angle=-90,
%  nodeanchor=east,
%  vshift=-5mm,
%  opacity=1,
%  scale=3,
%  contents=Draft
%}


\begin{document}

%\maketitle


%\thispagestyle{empty}
\begin{titlepage}
\vspace*{\stretch{1.0}}
\begin{center}

\Large{\textbf{Utilizing the Hamiltonian dynamics to study resonant interactions
of whistler-mode waves and electrons in the solar wind}} \\
\vspace{2.5cm}
\large{Tien Vo} \\
\vspace{2.5cm}
Submitted under the supervision of Professor Cynthia Cattell and Professor
Robert Lysak to the University Honors Program at the University of
Minnesota--Twin Cities in partial fulfillment of the requirements for the degree
of Bachelor of Science, summa cum laude in Physics. \\
\vspace{2.5cm}
\today

\end{center}
\vspace*{\stretch{2.0}}
\end{titlepage}

\newpage
\section*{Acknowledgements}

First and foremost, I thank Professor Cattell for having encouraged me to
explore the multiple aspects of this project. With her assistance and advice, I
have grown as a researcher from having the freedom to utilize my interests in
math, computation, and physics. Second, I thank Professor Lysak for his help in
the derivations of the once very intimidating math in this thesis. I
would also like to thank Dr. Vadim Roytershteyn for the insightful
conversations about PIC simulations, which greatly helped shape my code.
I thank Aaron West for his great contributions to the research. And finally, I thank my family for their love and support of my academic pursuit. This thesis owes its existence to their belief in me.

The Minnesota Supercomputing Institute (MSI) at the University of Minnesota
provided the resources that contributed to the research results reported with
this thesis. (URL: https://www.msi.umn.edu).


\newpage

\begin{abstract}
    The role of large amplitude whistler waves in the energization and
    scattering of solar wind electrons has long been an interesting problem in
    Space Physics. To study this wave-particle interaction, we developed a
    vectorized test particle simulation with a variational calculation of the
    Lyapunov exponents. From using secular perturbation theory on this
    Hamiltonian system of wave and particle, we confirmed that the pitch angle
    diffusion of the particle was along the constant Hamiltonian surface and
    that it was driven by the interaction with the resonance surfaces. We also
    showed that oblique whistlers could efficiently scatter field-aligned strahl
    electrons into the halo population in the solar wind. We demonstrated
    through simulation that these waves were capable of generating horn-like
    features in the velocity distribution function, similar to recent PIC
    simulation results in the literature.
\end{abstract}

\newpage

\singlespacing
\tableofcontents
\doublespacing


\newpage
\section{Introduction}\label{sec:intro}
\input{chapters/introduction/intro}

\section{Theory}\label{sec:theory}
\input{chapters/theory/whistler_fields}
\input{chapters/theory/particle_dynamics}
\input{chapters/theory/resonance_surfaces}

\section{Simulation}\label{sec:simulation}
\input{chapters/simulation/particle_advance}
\input{chapters/simulation/lyapunov_exponents}

\section{Diagnostics}\label{sec:diagnostics}
\input{chapters/diagnostics/simulation_parameters}
\input{chapters/diagnostics/lyapunov_exponents}

\section{Analysis}\label{sec:analysis}
\input{chapters/analysis/analysis}

\input{chapters/conclusion/conclusions}


\newpage
\appendix
\appendixpage

\input{chapters/appendix/theorems}
\input{chapters/appendix/hamiltonian_eom}
\input{chapters/appendix/H_R_surfaces}
\input{chapters/appendix/expansion_operator}
%\input{chapters/appendix/VDF_figures}


\newpage
\bibliographystyle{aasjournal}
\bibliography{references}

\end{document}
